---
title: "Selective Inference for K-means Clustering on Version A AML Segments"
author: "Junyi Hou"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```

## Section 1: Background and Objective

This document implements selective inference for k-means clustering on the Version A AML customer segmentation pipeline.

Our existing pipeline applies FAMD to mixed-type customer features and then uses k-means clustering to identify behavioral customer segments. These clusters are later used for interpretation, within-cluster anomaly detection, and downstream risk scoring.

However, classical post-clustering comparisons are biased because the same data are used both to create the clusters and to test whether the clusters differ. The KmeansInference package addresses this problem by providing selective p-values for testing whether two estimated k-means clusters have different mean vectors, while accounting for the clustering step itself.

We implement selective inference following Chen and Witten (2022), which provides valid post-clustering p-values by accounting for selection bias introduced during k-means clustering.

The main goal of this analysis is:

1. Determine whether selected k-means clusters are statistically distinguishable after selection correction;
2. Assess whether the current choice of k may be over-segmenting or under-segmenting the data;
3. Generates reliability score, this is being used as a part of rule-based scorer.

This file focuses on Version A clustering only:
- Business customers
- Individual customers

```{r, include=FALSE}
library(tidyverse)
library(KmeansInference)
library(knitr)
library(kableExtra)
# Please use the following code to download the KmeansInference Package
# require("devtools")
# devtools::install_github("yiqunchen/KmeansInference")
# library(KmeansInference)
```

```{r}
business_after_path  <- "VersionA_Business_AfterFAMD.csv"
business_before_path <- "VersionA_Business_BeforeFAMD.csv"

individual_after_path  <- "VersionA_Individual_AfterFAMD.csv"
individual_before_path <- "VersionA_Individual_BeforeFAMD.csv"

business_after  <- read.csv(business_after_path, check.names = FALSE)
business_before <- read.csv(business_before_path, check.names = FALSE)

individual_after  <- read.csv(individual_after_path, check.names = FALSE)
individual_before <- read.csv(individual_before_path, check.names = FALSE)
```

```{r}
# ---- Basic dimensions ----
cat("Business AfterFAMD dimensions: ", dim(business_after), "\n")
cat("Business BeforeFAMD dimensions:", dim(business_before), "\n\n")

cat("Individual AfterFAMD dimensions: ", dim(individual_after), "\n")
cat("Individual BeforeFAMD dimensions:", dim(individual_before), "\n")
```
```{r}
# ---- Preview columns ----
cat("Business AfterFAMD column names:\n")
print(colnames(business_after))

cat("\nBusiness BeforeFAMD first 15 columns:\n")
print(colnames(business_before)[1:min(15, ncol(business_before))])

cat("\nIndividual AfterFAMD column names:\n")
print(colnames(individual_after))

cat("\nIndividual BeforeFAMD first 15 columns:\n")
print(colnames(individual_before)[1:min(15, ncol(individual_before))])
```
```{r}
# ---- Convert AfterFAMD data to numeric matrices ----
# These are the matrices to be used for k-means and selective inference.

X_business <- business_after %>%
  mutate(across(everything(), as.numeric)) %>%
  as.matrix()

X_individual <- individual_after %>%
  mutate(across(everything(), as.numeric)) %>%
  as.matrix()

# Rename columns
colnames(X_business) <- paste0("FAMD_", seq_len(ncol(X_business)))
colnames(X_individual) <- paste0("FAMD_", seq_len(ncol(X_individual)))

# ---- Validate FAMD matrices ----
cat("Business matrix mode: ", mode(X_business), "\n")
cat("Business matrix dimensions: ", dim(X_business), "\n")
cat("Any missing values in business matrix? ", anyNA(X_business), "\n\n")

cat("Individual matrix mode: ", mode(X_individual), "\n")
cat("Individual matrix dimensions: ", dim(X_individual), "\n")
cat("Any missing values in individual matrix? ", anyNA(X_individual), "\n")

```

```{r}
# ---- Summary statistics of FAMD components ----
business_famd_summary <- as.data.frame(X_business) %>%
  summarise(across(everything(),
                   list(mean = mean, sd = sd, min = min, max = max),
                   .names = "{.col}_{.fn}"))

individual_famd_summary <- as.data.frame(X_individual) %>%
  summarise(across(everything(),
                   list(mean = mean, sd = sd, min = min, max = max),
                   .names = "{.col}_{.fn}"))

cat("Business FAMD component summary:\n")
print(business_famd_summary)

cat("\nIndividual FAMD component summary:\n")
print(individual_famd_summary)
```

```{r}
# Keep customer IDs for interpretation later, but do not include them in X.
business_ids <- business_before$customer_id
individual_ids <- individual_before$customer_id

business_meta <- business_before %>%
  select(customer_id, province, industry_code)

individual_meta <- individual_before %>%
  select(customer_id, province, occupation_code)

cat("Analysis-ready business matrix created: X_business\n")
cat("Analysis-ready individual matrix created: X_individual\n")
```
Kmeans clustering and Selective-inference analysis is now conducted on the `AfterFAMD` datasets. The `BeforeFAMD` files will be retained only for customer identifiers and later interpretation, but not used directly in the selective inference calculations.

```{r}
# Section 2. K-means clustering with fixed seed

k_business <- 6
k_individual <- 6
iter_max_value <- 30

fixed_seed <- 1234

cat("Using fixed seed:", fixed_seed, "\n")
```

```{r}
# 2.1 Run kmeans_estimation for clustering

business_est <- kmeans_estimation(
  X = X_business,
  k = k_business,
  iter.max = iter_max_value,
  seed = fixed_seed
)

individual_est <- kmeans_estimation(
  X = X_individual,
  k = k_individual,
  iter.max = iter_max_value,
  seed = fixed_seed
)

cat("kmeans_estimation completed.\n")
```

```{r}
# 2.2 Extract final clusters from kmeans_estimation

cluster_business <- business_est$cluster[[business_est$iter]]
cluster_individual <- individual_est$cluster[[individual_est$iter]]

cluster_business_raw <- paste0("business_", cluster_business)
cluster_individual_raw <- paste0("individual_", cluster_individual)

cat("Business clusters from kmeans_estimation:\n")
print(table(cluster_business))

cat("\nIndividual clusters from kmeans_estimation:\n")
print(table(cluster_individual))

```

```{r}
# 2.3 Cluster size summaries

business_cluster_sizes <- data.frame(
  cluster_num = sort(unique(cluster_business))
) %>%
  dplyr::mutate(
    cluster_name = paste0("business_", cluster_num),
    n = as.integer(table(cluster_business)[as.character(cluster_num)])
  ) %>%
  dplyr::arrange(cluster_num)

individual_cluster_sizes <- data.frame(
  cluster_num = sort(unique(cluster_individual))
) %>%
  dplyr::mutate(
    cluster_name = paste0("individual_", cluster_num),
    n = as.integer(table(cluster_individual)[as.character(cluster_num)])
  ) %>%
  dplyr::arrange(cluster_num)

print(business_cluster_sizes)
print(individual_cluster_sizes)

```

```{r}
# 2.4 Prepare for Analysis

business_analysis_df <- data.frame(
  customer_id = business_ids,
  cluster_raw = cluster_business_raw,
  cluster_num = cluster_business,
  X_business,
  check.names = FALSE
)

individual_analysis_df <- data.frame(
  customer_id = individual_ids,
  cluster_raw = cluster_individual_raw,
  cluster_num = cluster_individual,
  X_individual,
  check.names = FALSE
)
```

```{r}
# 2.5 Build all pair combinations

get_all_pairs <- function(clusters, prefix) {
  clusts <- sort(unique(clusters))
  
  pairs <- t(combn(clusts, 2))
  pairs_df <- data.frame(
    cluster_a = pairs[, 1],
    cluster_b = pairs[, 2],
    stringsAsFactors = FALSE
  ) %>%
    dplyr::mutate(
      cluster_a_name = paste0(prefix, "_", cluster_a),
      cluster_b_name = paste0(prefix, "_", cluster_b)
    )
  
  pairs_df
}

pairs_business <- get_all_pairs(cluster_business, "business")
pairs_individual <- get_all_pairs(cluster_individual, "individual")

print(pairs_business)
print(pairs_individual)
```

```{r}
# 3. Selective inference using package-owned clustering

# 3.1 Robust sigma estimator
estimate_MED <- function(X) {
  X_centered <- X
  for (j in seq_len(ncol(X_centered))) {
    X_centered[, j] <- X_centered[, j] - median(X_centered[, j])
  }
  sigma_hat <- sqrt(median(X_centered^2) / qchisq(0.5, df = 1))
  return(sigma_hat)
}

sig_business <- estimate_MED(X_business)
sig_individual <- estimate_MED(X_individual)

cat("Estimated sigma (business):", sig_business, "\n")
cat("Estimated sigma (individual):", sig_individual, "\n")
```



```{r}
# 3.2 Pairwise selective inference helper
# Note: After experimenting, each run of the entire rmd takes 30 minute to 2 hours, please try to close most other applications running on the computer when running this file.
run_pairwise_inference <- function(X, pairs_df, k, sig, seed, iter.max = 30) {
  out_list <- vector("list", nrow(pairs_df))
  start_total <- Sys.time()

  for (i in seq_len(nrow(pairs_df))) {
    iter_start <- Sys.time()

    c1 <- pairs_df$cluster_a[i]
    c2 <- pairs_df$cluster_b[i]

    cat("\n====================================\n")
    cat("Running pair", i, "of", nrow(pairs_df), "\n")
    cat("Clusters:", c1, "vs", c2, "\n")
    cat("Start time:", as.character(iter_start), "\n")
    flush.console()

    inf_obj <- kmeans_inference(
      X = X,
      k = k,
      cluster_1 = c1,
      cluster_2 = c2,
      sig = sig,
      iter.max = iter.max,
      seed = seed
    )

    inf_sum <- summary(inf_obj)

    iter_end <- Sys.time()
    iter_time <- as.numeric(difftime(iter_end, iter_start, units = "secs"))

    cat("Finished pair", i, "\n")
    cat("Time taken (seconds):", round(iter_time, 2), "\n")
    flush.console()

    out_list[[i]] <- data.frame(
      cluster_a = c1,
      cluster_b = c2,
      cluster_a_name = pairs_df$cluster_a_name[i],
      cluster_b_name = pairs_df$cluster_b_name[i],
      test_stat = inf_sum$test_stat,
      p_selective = inf_sum$p_selective,
      p_naive = inf_sum$p_naive,
      time_sec = iter_time,
      stringsAsFactors = FALSE
    )
  }

  total_time <- as.numeric(difftime(Sys.time(), start_total, units = "secs"))
  cat("\n====================================\n")
  cat("TOTAL TIME (seconds):", round(total_time, 2), "\n")
  cat("====================================\n")

  out <- do.call(rbind, out_list)
  out <- out[order(out$p_selective), ]
  rownames(out) <- NULL
  return(out)
}
```

```{r}
# 3.3 Run inference for both segments
results_list <- list(
  business = run_pairwise_inference(
    X = X_business,
    pairs_df = pairs_business,
    k = k_business,
    sig = sig_business,
    seed = fixed_seed,
    iter.max = iter_max_value
  ),
  individual = run_pairwise_inference(
    X = X_individual,
    pairs_df = pairs_individual,
    k = k_individual,
    sig = sig_individual,
    seed = fixed_seed,
    iter.max = iter_max_value
  )
)
```

```{r}
# 3.4 Add significance flags
results_list <- lapply(results_list, function(df) {
  df %>%
    dplyr::mutate(
      selective_sig_05 = p_selective < 0.05,
      naive_sig_05 = p_naive < 0.05
    )
})
```

```{r}
# 3.5 Extract final result tables
business_inference_results <- results_list$business
individual_inference_results <- results_list$individual

cat("Business selective inference results:\n")
print(business_inference_results)

cat("\nIndividual selective inference results:\n")
print(individual_inference_results)
```

```{r}
# 3.6 Save outputs
write.csv(
  business_inference_results,
  "business_selective_inference_results.csv",
  row.names = FALSE
)

write.csv(
  individual_inference_results,
  "individual_selective_inference_results.csv",
  row.names = FALSE
)

cat("Selective inference result files exported.\n")
```

```{r}
write.csv(
  business_analysis_df %>% dplyr::select(customer_id, cluster_raw, cluster_num),
  "business_kmeans_estimation_labels.csv",
  row.names = FALSE
)

write.csv(
  individual_analysis_df %>% dplyr::select(customer_id, cluster_raw, cluster_num),
  "individual_kmeans_estimation_labels.csv",
  row.names = FALSE
)
```

# Reliability Score:

```{r}
# 3.7 Reliability score calculation from pairwise selective inference results

calculate_reliability_scores <- function(inference_df, prefix, k_value) {
  
  # Keep only significant pairs
  sig_pairs <- inference_df %>%
    dplyr::filter(selective_sig_05)

  # Count how many significant pairs involve each cluster
  reliability_df <- data.frame(
    cluster_num = seq_len(k_value)
  ) %>%
    dplyr::rowwise() %>%
    dplyr::mutate(
      significant_pairs_involving_cluster = sum(
        sig_pairs$cluster_a == cluster_num | sig_pairs$cluster_b == cluster_num
      ),
      reliability_score = significant_pairs_involving_cluster / (k_value - 1),
      cluster_name = paste0(prefix, "_", cluster_num)
    ) %>%
    dplyr::ungroup() %>%
    dplyr::select(
      cluster_num,
      cluster_name,
      significant_pairs_involving_cluster,
      reliability_score
    ) %>%
    dplyr::arrange(cluster_num)

  return(reliability_df)
}

business_reliability_scores <- calculate_reliability_scores(
  inference_df = business_inference_results,
  prefix = "business",
  k_value = k_business
)

individual_reliability_scores <- calculate_reliability_scores(
  inference_df = individual_inference_results,
  prefix = "individual",
  k_value = k_individual
)

cat("Business reliability scores:\n")
print(business_reliability_scores)

cat("\nIndividual reliability scores:\n")
print(individual_reliability_scores)
```

```{r}
# 3.8 Save reliability score outputs
write.csv(
  business_reliability_scores,
  "business_cluster_reliability_scores.csv",
  row.names = FALSE
)

write.csv(
  individual_reliability_scores,
  "individual_cluster_reliability_scores.csv",
  row.names = FALSE
)

cat("Cluster reliability score files exported.\n")
```

## References

Chen, Y. T., & Witten, D. M. (2022). *Selective inference for k-means clustering*. 
https://doi.org/10.48550/arxiv.2203.15267